quollr: An R Package for Visalizing 2D Models in High Dimensional Space

### 2024-03-04 paper-quollr.Rmd
Add buffer parameter image
Added weigted mean cal

---
title: "quollr: An R Package for Visalizing 2D Models in High Dimensional Space"
abstract: >
  An abstract of less than 150 words.
draft: true
author:  
  - name: Jayani P.G. Lakshika
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    url: https://jayanilakshika.netlify.app/
    orcid: 0000-0002-6265-6481
    email:  \email{jayani.piyadigamage@monash.edu}
  - name: Dianne Cook
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    url: http://www.dicook.org/
    email: dicook@monash.edu
    orcid: 0000-0002-3813-7155
  - name: Paul Harrison
    affiliation: Monash University
    address: MGBP, BDInstitute, VIC 3800 Australia
    email: paul.harrison@monash.edu
    orcid: 0000-0002-3980-268X
  - name: Michael Lydeamore
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    email: michael.lydeamore@monash.edu
    orcid: 0000-0001-6515-827X
  - name: Thiyanga S. Talagala
    affiliation: University of Sri Jayewardenepura
    address: Department of Statistics, Gangodawila, Nugegoda 10100 Sri Lanka
    url: https://thiyanga.netlify.app/
    email: ttalagala@sjp.ac.lk
    orcid: 0000-0002-0656-9789
type: package
preamble: >
  \usepackage{amsmath}
  \usepackage{array}
output: 
  rjtools::rjournal_article:
    self_contained: yes
    toc: no
bibliography: RJreferences.bib

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)
source("quollr_code.R", local = TRUE)
source("nldr_code.R", local = TRUE)
```

```{r load-libraries}
#library(quollr)
library(knitr)
library(kableExtra)
library(readr)
library(ggplot2)
library(dplyr)
library(ggbeeswarm)
library(Rtsne)
library(umap)
library(phateR)
library(reticulate)
library(rsample)

set.seed(20230531)

# use_python("~/miniforge3/envs/pcamp_env/bin/python")
# use_condaenv("pcamp_env")
# 
# reticulate::source_python(paste0(here::here(), "/scripts/function_scripts/Fit_PacMAP_code.py"))
# reticulate::source_python(paste0(here::here(), "/scripts/function_scripts/Fit_TriMAP_code.py"))

@@

```{r}
#| echo: false

training_data <- read_rds(file = "data/s_curve_noise_training.rds")
test_data <- read_rds(file = "data/s_curve_noise_test.rds")

s_curve_noise <- read_rds(file = "data/s_curve_noise.rds")
s_curve_noise_umap <- read_rds(file = "data/s_curve_noise_umap.rds")
```

```{r}
library(scales)
s_curve_noise_umap$UMAP1 <- rescale(s_curve_noise_umap$UMAP1)

y_min <- -sqrt(3)/2
y_max <- sqrt(3)/2

# y_min <- 0
# y_max <- sqrt(3)/2

s_curve_noise_umap$UMAP2 <- ((s_curve_noise_umap$UMAP2 - min(s_curve_noise_umap$UMAP2))/(max(s_curve_noise_umap$UMAP2) - min(s_curve_noise_umap$UMAP2))) * (y_max - y_min)

# Define the desired size of the hexagons
hex_size <- 0.2

# Compute the horizontal spacing (dx) and vertical spacing (dy) between hexagon centers
# dx <- sqrt(3) * hex_size
# dy <- 1.5 * hex_size

# x_length <- 1
# y_length_scaled <- diff(range(s_curve_noise_umap$UMAP2))

# Compute the number of bins along the x-axis and y-axis
# num_bins_x <- ceiling(x_length / dx)
# num_bins_y <- ceiling(y_length_scaled / dy)
```

```{r}
# s_curve_noise_umap[,1:(NCOL(s_curve_noise_umap) - 1)] <- scale(s_curve_noise_umap[,1:(NCOL(s_curve_noise_umap) - 1)])
```

@@

```{r, eval=FALSE}
library(tools)
package_dependencies("quollr")
```

@@

```{r, echo=FALSE}
datasets_tb <- tibble(dt = c("s_curve_noise",
                             "s_curve_noise_training", 
                             "s_curve_noise_test", 
                             "s_curve_noise_umap"), 
                      text = c("Simulated 3D S-curve data with additional four noise dimensions.",
                               "Training data derived from S-curve data.", 
                               "Test data derived from S-curve data.", 
                               "UMAP 2D embedding data of S-curve data (n_neighbors: 15, min_dist: 0.1)."))
```

@@

@@

Hexagonal binning is a powerful technique for visualizing the density of data points in a 2-d space. Unlike traditional rectangular bins, hexagonal bins offer several advantages, including a more uniform representation of density and reduced visual bias. However, to effectively do the hexagonal binning, it's essential to compute the appropriate configurations based on the characteristics of the dataset.

Before computing hexagonal bin configurations, we need to determine the range of data along the x and y axes to establish the boundary for hexagonal binning. Additionally, choosing an appropriate hexagon size (the radius of the outer circle) is essential. By default, the `calculate_effective_x_bins()` and `calculate_effective_y_bins()` functions use a hexagon size of 1.07457. However, users can adjust the hexagon size to fit the data range and achieve regular hexagons without overlapping.

The hexagon size directly affects the number of bins generated. A higher hexagonal size will result in fewer bins, while a lower hexagonal size will lead to more bins. Therefore, there's always a trade-off depending on the dataset used. Users should consider their specific data characteristics when selecting the hexagon size.

```{r}
num_bins_x <- calculate_effective_x_bins(.data = s_curve_noise_umap,
                                         x = "UMAP1", hex_size = 0.2)
num_bins_x 
```

```{r}
num_bins_y <- calculate_effective_y_bins(.data = s_curve_noise_umap,
                                         y = "UMAP2", hex_size = 0.2)
num_bins_y 
```

### Generate full hex grid

Generating full hexagonal grid contains main three steps:

1. Generate all the hexagonal bin centroids

```{r}
all_centroids_df <- generate_full_grid_centroids(nldr_df = s_curve_noise_umap, 
                                                 x = "UMAP1", y = "UMAP2", 
                                                 num_bins_x = num_bins_x, 
                                                 num_bins_y = num_bins_y, 
                                                 x_start = NA, y_start = NA, 
                                                 hex_size = 0.2)

glimpse(all_centroids_df)
```

2. Generate hexagonal coordinates

```{r}
hex_grid <- gen_hex_coordinates(all_centroids_df, hex_size = 0.2)
glimpse(hex_grid)
```

```{r}
ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_point(aes(x = c_x, y = c_y), color = "red") +
  coord_fixed()
```

```{r}
ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
  coord_fixed()
```

5. Assign data into hexagons

- Compute distances between nldr coordinates and hex bin centroids

- Find the hexagonal centroid that have the minimum distance

```{r}
s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, hex_grid)
```

6. Compute standardized counts

- Compute number of data points within each hexagon

- Compute standardise count by dividing the counts by the maximum

```{r}
df_with_std_counts <- compute_std_counts(nldr_df_with_hex_id = s_curve_noise_umap_with_id)
```

7. Extract full grid info

- Assign standardize counts for hex bins

- Join with the hexagonal coordinates

```{r}
hex_full_count_df <- generate_full_grid_info(nldr_df = s_curve_noise_umap, hex_grid)
```

```{r}
ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_point(data = s_curve_noise_umap, aes(x = UMAP1, y = UMAP2), color = "blue") +
  coord_fixed()
```

```{r}
ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = id, fill = std_counts)) +
  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
  coord_fixed()

#### Buffer size

When generating hexagonal bins in R, a buffer is often included to ensure that the data points are evenly distributed within the bins and to prevent edge effects. The buffer helps in two main ways:

1. **Preventing Edge Effects**: Without a buffer, the outermost data points might fall near the boundary of the hexagonal grid, leading to incomplete bins or uneven distribution of data. By adding a buffer, you create a margin around the outer edges of the grid, ensuring that all data points are fully enclosed within the bins.

2. **Ensuring Even Distribution**: The buffer allows for a smoother transition between adjacent bins. This helps in cases where data points are not perfectly aligned with the grid lines, ensuring that each data point is assigned to the nearest bin without bias towards any specific direction.

Overall, including a buffer when generating hexagonal bins helps to produce more accurate and robust binning results, particularly when dealing with real-world data that may have irregular distributions or boundary effects.

### Construct the 2D model with different options

```{r}
df_bin_centroids <- hex_full_count_df[complete.cases(hex_full_count_df[["std_counts"]]), ] |>
  dplyr::select("c_x", "c_y", "hexID", "std_counts") |>
  dplyr::distinct() |>
  dplyr::rename(c("x" = "c_x", "y" = "c_y"))

df_bin_centroids
```

### Construct the high-D model with different options

```{r}
## To generate a data set with high-D and 2D training data
df_all <- training_data |> dplyr::select(-ID) |>
  dplyr::bind_cols(s_curve_noise_umap_with_id)

## To generate averaged high-D data

df_bin <- avg_highD_data(.data = df_all, column_start_text = "x") ## Need to pass ID column name
```

### Generate the triangular mesh

```{r}
tr1_object <- triangulate_bin_centroids(df_bin_centroids, x = "x", y = "y")
tr_from_to_df <- generate_edge_info(triangular_object = tr1_object)
```

### Compute parameter defaults

#### Shift the hexagonal grid origin

If shift_x happen to the positive direction of x it should input as a positive value, if not other way
If shift_y happen to the positive direction of y it should input as a positive value, if not other way

1. Assign shift along the x and y axis (limited the amount should less than the cell_diameter)

2. Generate bounds with shift origin

#### Benchmark value to remove the low-density hexagons

#### Benchmark value to remove the long edges

```{r}
## Compute 2D distances
distance <- cal_2d_dist(tr_from_to_df_coord = tr_from_to_df)

## To plot the distribution of distance
plot_dist <- function(distance_df){
  distance_df$group <- "1"
  dist_plot <- ggplot(distance_df, aes(x = group, y = distance)) +
    geom_quasirandom()+
    ylim(0, max(unlist(distance_df$distance))+ 0.5) + coord_flip()
  return(dist_plot)
}

plot_dist(distance)

benchmark <- find_benchmark_value(distance_edges = distance, distance_col = "distance")
benchmark

benchmark <- 0.75
```

### Model function

`fit_high_d_model()` function is used to generate the 2D model and the high-D model.

```{r}
## To generate a data set with high-D and 2D training data
df_all <- training_data |> dplyr::select(-ID) |>
  dplyr::bind_cols(s_curve_noise_umap_with_id)

## To generate averaged high-D data

df_bin <- avg_highD_data(.data = df_all, column_start_text = "x") ## Need to pass ID column name
```

@@

To predict the 2D embeddings for a new data point using the trained high-D model, we follow a series of steps outlined below:

1. **Compute high-dimensional Euclidean distance**: Calculate the Euclidean distance between each new data point and the lift model's coordinates in the high-dimensional space. This distance metric helps identify the nearest lift model coordinates to each new data point.

2. **Find the nearest lift model coordinates**: Determine the lift model's high-dimensional coordinates that are closest to each new data point based on the computed distances. This step helps establish a correspondence between the new data points and the lift model's representation in the high-dimensional space.

3. **Map the hexagonal bin ID**: Assign each new data point to a hexagonal bin based on its nearest lift model coordinates. This mapping ensures that each data point is associated with a specific region in the hexagonal grid representation of the high-dimensional space.

4. **Map the hexagonal bin centroid coordinates in 2D**: Transform the hexagonal bin centroid coordinates from the high-dimensional space to the 2D embedding space using the trained lift model. This mapping provides the 2D embeddings for the new data points, allowing them to be visualized and analyzed in a lower-dimensional space.

The `predict_2d_embeddings()` function is used to predict 2D embeddings for a new data point. The inputs for the function are test data, bin centroid coordinates in 2D, lifting model coordinates in high-D, and the type of the NLDR technique. The output contains the predicted 2D embeddings along with the predicted hexagonal id and the ID of the test data.

```{r}
pred_df_test <- predict_2d_embeddings(test_data = training_data,
df_bin_centroids = df_bin_centroids, df_bin = df_bin, type_NLDR = "UMAP")

glimpse(pred_df_test)
```

### Making summaries

There are two important summaries that should be made when constructing the model on a dataset using a specific NLDR technique, the Mean Square Error (MSE) and Akaike Information Criterion (AIC) which measure prediction accuracy. The function that perform the summaries called `generate_eval_df()`. The output of this function is a list which contains MSE and AIC. The code below uses `generate_eval_df()` to find the MSE and the AIC values for the model constructed on UMAP 2D embeddings generated from the `s_curve_noise` training data.

```{r}
generate_summary(test_data = training_data, prediction_df = pred_df_test,
                 df_bin = df_bin, col_start = "x")
```

### Visualizations

@@

@@

The `show_langevitour()` function enables dynamic visualization of the 2D model alongside the high-dimensional (high-D) data in its original space. This visualization is facilitated by langevitour object, allowing users to interactively explore the relationship between the 2D embeddings and the underlying high-dimensional data. The main arguments are shown in Table

```{r}
trimesh <- ggplot(df_bin_centroids, aes(x = x, y = y)) +
  geom_point(size = 0.1) +
  geom_trimesh() +
  coord_equal()

trimesh
```

```{r}
trimesh_gr <- colour_long_edges(distance_edges = distance, benchmark_value = benchmark,
                                tr_from_to_df_coord = tr_from_to_df, distance_col = "distance")

trimesh_gr
```

```{r}
trimesh_removed <- remove_long_edges(distance_edges = distance, benchmark_value = benchmark,
                                     tr_from_to_df_coord = tr_from_to_df, distance_col = "distance")
trimesh_removed
```

```{r}
tour1 <- show_langevitour(df_all, df_bin, df_bin_centroids, 
                          benchmark_value = benchmark,
                          distance = distance, distance_col = "distance", 
                          use_default_benchmark_val = FALSE, 
                          column_start_text = "x")
tour1
```

### Find non-empty bins

@@

This article is created using \CRANpkg{knitr} [@knitr] and \CRANpkg{rmarkdown} [@rmarkdown] in R with the `rjtools::rjournal_article` template. The source code for reproducing this paper can be found at: [https://github.com/JayaniLakshika/paper-quollr](https://github.com/JayaniLakshika/paper-quollr).

>>>>

---
title: "quollr: An R Package for Visalizing 2D Models in High Dimensional Space"
abstract: >
  An abstract of less than 150 words.
draft: true
author:  
  - name: Jayani P.G. Lakshika
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    url: https://jayanilakshika.netlify.app/
    orcid: 0000-0002-6265-6481
    email:  \email{jayani.piyadigamage@monash.edu}
  - name: Dianne Cook
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    url: http://www.dicook.org/
    email: dicook@monash.edu
    orcid: 0000-0002-3813-7155
  - name: Paul Harrison
    affiliation: Monash University
    address: MGBP, BDInstitute, VIC 3800 Australia
    email: paul.harrison@monash.edu
    orcid: 0000-0002-3980-268X
  - name: Michael Lydeamore
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    email: michael.lydeamore@monash.edu
    orcid: 0000-0001-6515-827X
  - name: Thiyanga S. Talagala
    affiliation: University of Sri Jayewardenepura
    address: Department of Statistics, Gangodawila, Nugegoda 10100 Sri Lanka
    url: https://thiyanga.netlify.app/
    email: ttalagala@sjp.ac.lk
    orcid: 0000-0002-0656-9789
type: package
preamble: >
  \usepackage{amsmath}
  \usepackage{array}
output: 
  rjtools::rjournal_article:
    self_contained: yes
    toc: no
bibliography: RJreferences.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)

@@

```{r load-libraries}
library(quollr)
library(tibble)
library(knitr)
library(kableExtra)
library(ggplot2)

set.seed(20230531)
```

@@

@@

```{r, echo=FALSE}
datasets_tb <- tibble(dt = c("s_curve_noise",
                             "s_curve_noise_training", 
                             "s_curve_noise_test", 
                             "s_curve_noise_umap",
                             "s_curve_noise_umap_predict",
                             "s_curve_noise_umap_scaled"), 
                      text = c("Simulated 3D S-curve data with additional four noise dimensions.",
                               "Training data derived from S-curve data.", 
                               "Test data derived from S-curve data.", 
                               "UMAP 2D embedding data of S-curve data (n_neighbors: 15, min_dist: 0.1).",
                               "Predicted UMAP 2D embedding data of S-curve data",
                               "Scaled UMAP 2D embedding data of S-curve data"))
```

@@

## Preprocessing

In the preprocessing stage, we aim to prepare our data to fit within the bounds required for regular hexagonal binning, ensuring effective visualization. To achieve this, we implement two key scaling steps. Firstly, we scale the first 2D embedding component to range between $0$ and $1$, ensuring that all data points fall within this normalized interval. Secondly, we scale the second 2D embedding component to range between $0$ and $y_{max}$. This adjustment helps maintain the necessary symmetry and spacing required for regular hexagons.

```{=tex}
\begin{equation}\label{eq:aspect_ratio}
ar = \frac{r_2}{r_1} \tag{1}
\end{equation}
```
```{=tex}
\begin{equation}\label{eq:y_max}
y_{max} = ceiling(\frac{ar}{hr}) * hr \tag{2}
\end{equation}
```

```{r}
scaled_data <- gen_scaled_data(data = s_curve_noise_umap, x = "UMAP1", 
                y = "UMAP2", hex_ratio = NA)
glimpse(scaled_data)
```

## Construct the 2D model

@@

## Construct the high-D model

## Model function

The `fit_highd_model()` function is used to generate both the 2D and high-D models.

```{r}
fit_highd_model(training_data = s_curve_noise_training, x = "UMAP1", y = "UMAP2",
nldr_df_with_id = s_curve_noise_umap_scaled, col_start_2d = "UMAP", col_start_highd = "x")
```

## Model summaries

@@

There are some of NLDR techniques that don't give any functions for predictions. In that sense our methodology facilitates to compute 2D embedding with our workflow.

### Goodness of fit statistics

There are two Goodness of fit statistics were produced. MSE and AIC which interpret the model accuracy.

## Visualizations

@@

`r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:lgvis-tb-html)', '\\@ref(tab:lgvis-tb-pdf)'))`

```{r, echo=FALSE}
lg_vis_tb <- tibble(dt = c(".data", 
                           "benchmark_value", 
                           "triangular_object", 
                           "distance_col"), 
                    text = c("The data frame containing the edge information.",
                             "The threshold value to determine long edges.", 
                             "The triangular object containing the mesh information.", 
                             "The column name in `.data` representing the distances."))
```

```{r lgvis-tb-html, eval=is_html_output(), echo=FALSE}
lg_vis_tb |> 
  kable(caption = "The main arguments for `vis_lg_mesh()` and `vis_rmlg_mesh()`", col.names = c("argument", "explanation")) 
```

```{r lgvis-tb-pdf, eval=is_latex_output(), echo=FALSE}
lg_vis_tb |>
  kable(caption = "The main arguments for `vis\\_lg\\_mesh()` and `vis\\_rmlg\\_mesh()`", format="latex", col.names = c("argument", "explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
```

@@

The `show_langevitour()` function enables dynamic visualization of the 2D model alongside the high-dimensional (high-D) data in its original space. This visualization is facilitated by langevitour object, allowing users to interactively explore the relationship between the 2D embeddings and the underlying high-dimensional data. The main arguments are shown in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:dyvis-tb-html)', '\\@ref(tab:dyvis-tb-pdf)'))`.

```{r, echo=FALSE}
dyn_vis_tb <- tibble(dt = c("df", 
                            "df_b", 
                            "df_b_with_center_data", 
                            "benchmark_value",
                            "distance_df",
                            "distance_col",
                            "use_default_benchmark_val",
                            "column_start_text"), 
                     text = c("A data frame containing the high-dimensional data.",
                              "A data frame containing the high-dimensional coordinates of bin centroids/means.",
                              "The dataset with hexbin centroids/ means.", 
                              "The benchmark value used to remove long edges (optional).", 
                              "The distance dataframe.",
                              "The name of the distance column.",
                              "Logical, indicating whether to use default benchmark value  to remove long edges(default is FALSE).",
                              "The text that begin the column name of the high-dimensional data."))
```

```{r dyvis-tb-html, eval=is_html_output(), echo=FALSE}
dyn_vis_tb |> 
  kable(caption = "The main arguments for `show_langevitour()`", col.names = c("argument", "explanation")) 
```

```{r dyvis-tb-pdf, eval=is_latex_output(), echo=FALSE}
dyn_vis_tb |>
  kable(caption = "The main arguments for `show\\_langevitour()`", format="latex", col.names = c("argument", "explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
```

## Distance with bin means

```{r, echo=FALSE}
compute_weights <- function(nldr_df_with_hex_id) {

  ## To get the 2D embeddings average of each bin
  bin_val_hexagons <- nldr_df_with_hex_id |>
    dplyr::group_by(hb_id) |>
    dplyr::summarise(dplyr::across(tidyselect::everything(), mean))

  ## Rename columns of averaged 2D embeddings
  names(bin_val_hexagons) <- c("hb_id", paste0("avg_", tolower(names(nldr_df_with_hex_id)[1:2])))

  ## To calculate distances from average point
  nldr_with_avg_all <- dplyr::inner_join(bin_val_hexagons , nldr_df_with_hex_id,
                                         by = c("hb_id" = "hb_id"))

  col_names <- c(names(bin_val_hexagons), names(nldr_df_with_hex_id)[1:2], "distance")

  ## Initialize the vectors to store data
  hexids <- integer(0)
  emb1_vec <- numeric(0)
  emb2_vec <- numeric(0)
  weight_vec <- numeric(0)

  for(hb_id in unique(nldr_with_avg_all$hb_id)){

    ## These are the weights for weighted mean
    weighted_mean_df <- nldr_with_avg_all |>
      dplyr::filter(hb_id == hb_id) |>
      cal_2d_dist(start_x = col_names[2], start_y = col_names[3], end_x = col_names[4],
                  end_y = col_names[5], select_vars = col_names)

    hexids <- c(hexids, weighted_mean_df$hb_id)
    emb1_vec <- c(emb1_vec, weighted_mean_df[[rlang::as_string(rlang::sym(col_names[4]))]])
    emb2_vec <- c(emb2_vec, weighted_mean_df[[rlang::as_string(rlang::sym(col_names[5]))]])
    weight_vec <- c(weight_vec, 1/ (weighted_mean_df$distance + 0.05))

  }

  weight_list <- list(hb_id = hexids, emb1_vec = emb1_vec, emb2_vec = emb2_vec,
                      weight_vec = weight_vec)
  names(weight_list) <- c("hb_id", names(nldr_df_with_hex_id)[1:2], "weights")

  return(weight_list)

}
```

```{r, echo=FALSE}
weighted_highD_data <- function(training_data, nldr_df_with_hex_id,
                                column_start_text = "x") {
  ## Remove ID column from training data
  training_data <- training_data |>
    dplyr::select(-ID)

  ## Join training data with 2D embeddings
  df_all <- dplyr::bind_cols(training_data, nldr_df_with_hex_id)

  ## To obtain weights corresponding to 2D embeddings
  weight_df <- as.data.frame(do.call(cbind,
                                     compute_weights(nldr_df_with_hex_id = nldr_df_with_hex_id)))

  ## Initialize the columns that need to use for joining
  joined_col_names <- names(nldr_df_with_hex_id)[1:2]

  ## To join the training data, 2D embeddings and weights
  weighted_mean_all <- dplyr::inner_join(df_all, weight_df,
                                         by = c("hb_id" = "hb_id",
                                                stats::setNames(joined_col_names,
                                                                joined_col_names)))
  ## List to store weighted means
  weighted_mean_df_list <- list()

  for (j in 1:NCOL(training_data)) {

    ## To compute weighted mean across all high-D coordinates
    weighted_mean_df_list[[j]] <- weighted_mean_all |>
      dplyr::select(hb_id, names(training_data)[j], weights) |>
      dplyr::group_by(hb_id) |>
      dplyr::summarise(dplyr::across(names(training_data)[j], ~ weighted.mean(., weights)))

  }

  ## To combine the elements given in the list
  weighted_mean <- Reduce(function(dtf1,dtf2) dplyr::full_join(dtf1,dtf2,by="hb_id"),
                          weighted_mean_df_list)

  ## Column names start with x
  weighted_mean <- weighted_mean |>
    dplyr::select(hb_id, tidyselect::starts_with(column_start_text))

  return(weighted_mean)
}
```

```{r, echo=FALSE}

extract_hexbin_mean <- function(nldr_df_with_hex_id, counts_df) {

  ## To arrange the hexagon IDs
  counts_df <- counts_df |>
    dplyr::arrange(hb_id)

  ## To compute hexagonal bin means
  hex_mean_df <- nldr_df_with_hex_id |>
    dplyr::group_by(hb_id) |>
    dplyr::summarise(dplyr::across(tidyselect::everything(), mean)) |>
    dplyr::arrange(hb_id) |>
    dplyr::filter(hb_id %in% counts_df$hb_id) |>
    dplyr::mutate(std_counts = counts_df$std_counts)

  ## Rename columns
  names(hex_mean_df) <- c("hexID", "c_x", "c_y", "std_counts")

  return(hex_mean_df)
}
```

```{r}
bin_list <- calc_bins(data = s_curve_noise_umap_scaled, 
                      x = "UMAP1", y = "UMAP2", 
                      hex_size = NA, buffer_x = NA, 
                      buffer_y = NA)
num_bins_x <- bin_list$num_x
num_bins_y <- bin_list$num_y

hb_obj <- hex_binning(data = s_curve_noise_umap_scaled, 
                      x = "UMAP1", y = "UMAP2", 
                      num_bins_x = num_bins_x, num_bins_y = num_bins_y, 
                      x_start = NA, y_start = NA, 
                      buffer_x = NA, buffer_y = NA, 
                      hex_size = NA, col_start = "UMAP")

all_centroids_df <- as.data.frame(do.call(cbind, hb_obj$centroids))
counts_df <- as.data.frame(do.call(cbind, hb_obj$std_cts))
nldr_df_with_hex_id <- as.data.frame(do.call(cbind, hb_obj$data_hb_id))

## To obtain bin centroids
df_bin_centroids <- extract_hexbin_mean(nldr_df_with_hex_id = nldr_df_with_hex_id,
                                             counts_df = counts_df)

df_all <- dplyr::bind_cols(s_curve_noise_training |> dplyr::select(-ID), nldr_df_with_hex_id)

ggplot() + 
  geom_trimesh(data = df_bin_centroids, mapping = aes(x = c_x, y = c_y)) +
  coord_fixed()

tr1_object <- tri_bin_centroids(hex_df = df_bin_centroids, x = "c_x", y = "c_y")
tr_from_to_df <- gen_edges(tri_object = tr1_object)
distance_df <- cal_2d_dist(tr_coord_df = tr_from_to_df, 
                           start_x = "x_from", start_y = "y_from", 
                           end_x = "x_to", end_y = "y_to", 
                           select_vars = c("from", "to", "distance"))

## averaged high-D data
df_bin <- weighted_highD_data(training_data = s_curve_noise_training, nldr_df_with_hex_id = nldr_df_with_hex_id, column_start_text = "x")

vis_lg_mesh(distance_edges = distance_df, benchmark_value = 1,
tr_coord_df = tr_from_to_df, distance_col = "distance")

vis_rmlg_mesh(distance_edges = distance_df, benchmark_value = 1,
tr_coord_df = tr_from_to_df, distance_col = "distance")

show_langevitour(df = df_all, df_b = df_bin, 
                 df_b_with_center_data = df_bin_centroids, 
                 benchmark_value = 1, distance = distance_df, 
                 distance_col = "distance", 
                 use_default_benchmark_val = FALSE, col_start = "x")
```

@@

This article is created using \CRANpkg{knitr} [@knitr] and \CRANpkg{rmarkdown} [@rmarkdown] in R with the `rjtools::rjournal_article` template. The source code for reproducing this paper can be found at: <https://github.com/JayaniLakshika/paper-quollr>.
